\fchapter{ACKNOWLEDGMENTS}
\pagestyle{plain}

First, I would like to thank Dr. Marie desJardins, my advisor, for guiding me through graduate school in many ways.
This work would not have been possible without her direction.
Specifically, I would like to thank her for dedicating an exceptional amount of personal time with her students and helping us through every step of the process.

I particularly thank Mr. John Way, my high school teacher at Sherwood High School, with whom I took three programming classes.
He was the teacher who not only introduced me to computer science, but also was the first teacher that cultivated my love for a subject and motivated me to succeed in something academic.
By inspiring me to excel in computer science back in 1999, he set me off on one of the most delightful journeys of my life.

Dr. Richard Chang introduced me to my first research experiences and motivated me to tackle hard problems.
Since taking algorithms with him in my undergraduate studies and completing my undergraduate thesis with him as my advisor, Dr. Chang has been a great friend and a great person to go to for advice.

Dr. Tim Oates has been a great help in graduate school.
His advice on machine learning topics is always clear, enlightening and interesting.
I appreciate all the time he has taken in helping guide my research---I was lucky to have him as a resource.

Many senior graduate students helped me as a young graduate student and served as great role models: Dr. Adam Anthony, Dr. Eric Eaton,  Dr. Blazej Bulka, and Dr. Tom Armstrong.
I thank Marc Picket I, who has been a good friend that has accompanied me throughout my entire graduate school experience.
He is always willing to listen to new ideas and give valuable feedback.
Many of the concepts and algorithms in this dissertation were developed on the dry-erase board over Marc's desk.
Traveling to conferences with Marc were some of the most enjoyable and unique experiences in graduate school.
The graduate student community at UMBC is a wonderful support group.
I would like to thank graduate students and post-docs Dr. Jesus Caban, Wes Griffin, Yasaman Haghpanah, Dr. Josh Jones, Niels Kasch, James MacGlashan, JC Montminy, Sourav Mukherjee, Patti Ordonez, Soumi Ray and Brandon Wilson for their support.

Dr.~{}Bill Rand and Forrest Stonedahl introduced me to NetLogo and much of the agent-based modeling terminology used in this dissertation.
My research was not focused on any particular type of system before meeting them.
My work naturally found its home in ABMs and NetLogo, which overall made my research far better than what it would have been otherwise.

I was assisted by undergraduate researchers in various stages of my research.
Thanks to Peter Hamilton, Kevin Winner, Doug Still and Nathaniel Wise for their contributions.

Finally, I thank all of my family for their love and support.


