\begin{titlepage}

\begin{center}
\vspace{0.1in}
\large{\bf Curriculum Vitae}
\bigskip \bigskip
\end{center}

\begin{flushleft}
  {\bf Name:}{\hspace{3mm}}Donald Miner\\
%	{\bf Permanent Address:}{\hspace{3mm}} MY-FULL-ADDRESS. \\
	{\bf Degree and date to be conferred:}{\hspace{3mm}}Ph.D. in Computer Science, May 2010. \\
%	{\bf Date of Birth:}{\hspace{3mm}}MY-BIRTHDATE. \\
%	{\bf Place of Birth:}{\hspace{3mm}}MY-PLACE-OF-BIRTH. \\
%	{\bf Secondary Education:}{\hspace{3mm}} MY-HIGH-SCHOOL, MY-HIGH-SCHOOLS-CITY, MY-HIGH-SCHOOLS-STATE.\\
	{\bf Collegiate institutions attended:}\\
	\begin{singlespace} 
	{\hspace{0.4in}}University of Maryland Baltimore County, Ph.D. Computer Science, 2006--2010. \\
	{\hspace{0.4in}}University of Maryland Baltimore County, B.S. Computer Science, 2002--2006. \\
	\end{singlespace} 
	\vspace{8pt}
	{\bf Professional publications:}\\
	\begin{singlespace} 

  {\hspace{0.4in}} \parbox[t]{5.5in}{ Don Miner, Marie desJardins. ``Predicting and Controlling System-Level Parameters of Multi-Agent Systems." In {\em Proceedings of the AAAI Fall Symposium on Complex Adaptive Systems and the Threshold Effect}, 2009.}\\
  {\vspace{5pt}}
  {\hspace{0.4in}} \parbox[t]{5.5in}{ Don Miner, Marie desJardins. ``Learning Non-Explicit Control Parameters of Self-Organizing Systems" (Extended Abstract). In {\em Proceedings of  the Third IEEE International Conference on Self-Adaptive and Self-Organizing Systems}, 2009.}\\
  {\vspace{5pt}}
  {\hspace{0.4in}} \parbox[t]{5.5in}{ Kevin Winner, Don Miner and Marie desJardins. ``Controlling Particle Swarm Optimization with Learned Parameters" (Extended Abstract). In {\em Proceedings of the Third IEEE International Conference on Self-Adaptive and Self-Organizing Systems}, 2009.}\\
  {\vspace{5pt}}
  {\hspace{0.4in}} \parbox[t]{5.5in}{ Don Miner, Marc Pickett, and Marie desJardins. ``Understanding the Brain's Emergent Properties." In {\em Proceedings of the Second Conference on Artificial General Intelligence}, 2009.}\\
  {\vspace{5pt}}
  {\hspace{0.4in}} \parbox[t]{5.5in}{ Marc Pickett, Don Miner, and Tim Oates, ``Essential Phenomena of General Intelligence," In {\em Proceedings of the First Conference on Artificial General Intelligence}, 2008.}\\
  {\vspace{5pt}}
  {\hspace{0.4in}} \parbox[t]{5.5in}{ Marc Pickett and Don Miner, ``Representation Change in The Marchitecture," In {\em Working Notes of the AAAI Fall Symposium on Representation Change}, 2007.}\\
  {\vspace{5pt}}
  {\hspace{0.4in}} \parbox[t]{5.5in}{ Marc Pickett, Don Miner, and Tim Oates, ``A Gauntlet for Evaluating Cognitive Architectures," In {\em Working Notes of the AAAI Workshop on Evaluating Architectures for Intelligence}, 2007. }\\
  {\vspace{5pt}}
  {\hspace{0.4in}} \parbox[t]{5.5in}{ Don Miner and James Athey, ``FCGlob: A New SELinux File Context Syntax," In {\em Proceedings of the Third Annual SELinux Symposium}, 2007.}\\
  {\vspace{5pt}}
   {\hspace{0.4in}} \parbox[t]{5.5in}{ James Athey, Christopher Ashworth, Frank Mayer and Don Miner, ``Towards Intuitive Tools for Managing SELinux: Hiding the Details but Retaining the Power," In {\em Proceedings of the Third Annual SELinux Symposium}, 2007.}\\
  {\vspace{5pt}}

  \end{singlespace} 

\pagebreak


	{\bf Professional positions held:}\\
	\begin{singlespace}
	{\hspace{0.4in}}\parbox[t]{5.5in}{Instructor -- University of Maryland, Baltimore County. (August 2008 -- Current).}\\
	{\vspace{5pt}}
	{\hspace{0.4in}}\parbox[t]{5.5in}{Research Assistant -- MAPLE Lab  (August 2007 -- May 2010).}\\
	{\vspace{5pt}}
	{\hspace{0.4in}}\parbox[t]{5.5in}{Software Developer in Test Intern -- Microsoft  (May 2007 -- August 2007).}\\
	{\vspace{5pt}}
	{\hspace{0.4in}}\parbox[t]{5.5in}{Teaching Assistant -- University of Maryland, Baltimore County  (August 2006 -- May 2007).}\\
	{\vspace{5pt}}
	{\hspace{0.4in}}\parbox[t]{5.5in}{Software Developer -- Tresys Technology  (May 2005 -- August 2006).}\\

	\end{singlespace}
\end{flushleft}
       
\end{titlepage}
\par\vfil


